---
title: "Week 4 Lab"
author: "Audrey Omidsalar"
date: "9/17/2021"
output:
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library('leaflet')
library('lubridate')
library('data.table')
library('tidyverse')
library('ggplot2')
```

## 1. Read in the Data

```{r cache=TRUE}
if (!file.exists("met_all.gz"))
  download.file(
    url = "https://raw.githubusercontent.com/USCbiostats/data-science-data/master/02_met/met_all.gz",
    destfile = "met_all.gz",
    method   = "libcurl",
    timeout  = 60
    )
met <- data.table::fread("met_all.gz")
```

## 2. Prepare the data

#### Remove temperatures less than -17C

```{r filter temps}
met <- met[temp>=-17]
summary(met)
```
#### After checking the data we should make the appropriate modifications.
```{r check for NA, include=FALSE}
met[,table(is.na(temp))]
met[,table(is.na(rh))]
met[,table(is.na(wind.sp))]
met[,table(is.na(vis.dist))]
met[,table(is.na(dew.point))]
met[,table(is.na(lat))]
met[,table(is.na(lon))]
met[,table(is.na(elev))]

```
```{r, include=FALSE}
met[,range(temp)]
met[,range(rh, na.rm = TRUE)]
met[,range(wind.sp, na.rm=TRUE)]
met[,range(vis.dist, na.rm=TRUE)]
met[,range(dew.point, na.rm=TRUE)]
met[,range(lat, na.rm=TRUE)]
met[,range(lon, na.rm=TRUE)]
met[,range(elev, na.rm=TRUE)]
```
Replace 9999 in elevation with NA
```{r replace with NA}
met[elev == 9999.0, elev := NA]
```

#### Generate a date variable
```{r date}
met[, ymd := as.Date(paste(year, month, day, sep = "-"))]
```

#### Using the data.table::week function, keep the observations of the first week of the month.

This outputs the week numbers according to ymd. For example, week 31 is the first week in August
```{r}
met[,table(week(ymd))]
```
#### Keep only the first week of the month, week=31
```{r first week}
met <- met[week(ymd) == 31]
```
#### Compute the mean by station of the variables temp, rh, wind.sp, vis.dist, dew.point, lat, lon, and elev. Creating a new datatable met_avg that stores these values.
```{r mean}
met_avg <- met[, .(
  temp     = mean(temp, na.rm=TRUE),
  rh       = mean(rh, na.rm=TRUE),
  wind.sp  = mean(wind.sp, na.rm=TRUE),
  vis.dist = mean(vis.dist, na.rm=TRUE),
  dew.point= mean(dew.point, na.rm=TRUE),
  lat      = mean(lat, na.rm=TRUE),
  lon      = mean(lon, na.rm=TRUE),
  elev     = mean(elev, na.rm=TRUE), USAFID
), by = "USAFID"]
```
#### Create a region variable for Northwest, Southwest, Northeast, Southeast based on lon = -98.00 and lat = 39.71 degrees. Count number of observations of each region variable.
```{r, cache=TRUE}
met_avg[lat >= 39.71 & lon <= -98, region := "Northwest"]
met_avg[lat < 39.71 & lon <= -98, region := "Southwest"]
met_avg[lat >= 39.71 & lon > -98, region := "Northeast"]
met_avg[lat < 39.71 & lon > -98, region := "Southeast"]
met_avg[, table(region, useNA="always")]
```
#### Create a categorical variable for elevation
```{r, cache=TRUE}
met_avg[,elev_cat := fifelse(elev > 252, "high", "low")]
```

