---
title: "Week 4 Lab"
author: "Audrey Omidsalar"
date: "9/17/2021"
output:
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library('leaflet')
library('lubridate')
library('data.table')
library('tidyverse')
library('ggplot2')
library('ggpubr')
```

## 1. Read in the Data

```{r cache=TRUE}
if (!file.exists("met_all.gz"))
  download.file(
    url = "https://raw.githubusercontent.com/USCbiostats/data-science-data/master/02_met/met_all.gz",
    destfile = "met_all.gz",
    method   = "libcurl",
    timeout  = 60
    )
met <- data.table::fread("met_all.gz")
```

## 2. Prepare the data

#### Remove temperatures less than -17C

```{r filter temps}
met <- met[temp>=-17]
summary(met)
```
#### After checking the data we should make the appropriate modifications.
```{r check for NA, include=FALSE}
met[,table(is.na(temp))]
met[,table(is.na(rh))]
met[,table(is.na(wind.sp))]
met[,table(is.na(vis.dist))]
met[,table(is.na(dew.point))]
met[,table(is.na(lat))]
met[,table(is.na(lon))]
met[,table(is.na(elev))]

```
```{r, include=FALSE}
met[,range(temp)]
met[,range(rh, na.rm = TRUE)]
met[,range(wind.sp, na.rm=TRUE)]
met[,range(vis.dist, na.rm=TRUE)]
met[,range(dew.point, na.rm=TRUE)]
met[,range(lat, na.rm=TRUE)]
met[,range(lon, na.rm=TRUE)]
met[,range(elev, na.rm=TRUE)]
```
Replace 9999 in elevation with NA
```{r replace with NA}
met[elev == 9999.0, elev := NA]
```

#### Generate a date variable using the as.Date() function

```{r date}
met[, ymd := as.Date(paste(year, month, day, sep = "-"))]
met
```

#### Using the data.table::week function, keep the observations of the first week of the month.

This outputs the week numbers according to the column with the date variable, ymd, set earlier. Week 31 is the first week of August.
```{r}
met[,table(week(ymd))]
```

#### Subset the dataset to keep only the first week of the month.

```{r first week}
met <- met[week(ymd) == 31]
met
```

#### Compute the mean by station of the variables temp, rh, wind.sp, vis.dist, dew.point, lat, lon, and elev. Creating a new datatable met_avg that stores these values.

```{r mean}
met_avg <- met[, .(
  temp     = mean(temp, na.rm=TRUE),
  rh       = mean(rh, na.rm=TRUE),
  wind.sp  = mean(wind.sp, na.rm=TRUE),
  vis.dist = mean(vis.dist, na.rm=TRUE),
  dew.point= mean(dew.point, na.rm=TRUE),
  lat      = mean(lat, na.rm=TRUE),
  lon      = mean(lon, na.rm=TRUE),
  elev     = mean(elev, na.rm=TRUE), USAFID
), by = "USAFID"]
met_avg
```

#### Create a region variable for Northwest, Southwest, Northeast, Southeast based on lon = -98.00 and lat = 39.71 degrees. Count number of observations of each region variable.

```{r, cache=TRUE}
met_avg[lat >= 39.71 & lon <= -98, region := "Northwest"]
met_avg[lat < 39.71 & lon <= -98, region := "Southwest"]
met_avg[lat >= 39.71 & lon > -98, region := "Northeast"]
met_avg[lat < 39.71 & lon > -98, region := "Southeast"]
met_avg[, table(region, useNA="always")]
```

#### Create a categorical variable for elevation

```{r, cache=TRUE}
met_avg[,elev_cat := fifelse(elev > 252, "high", "low")]
##Deleting extra "USAFID" column 1
met_avg <- met_avg[, -1]
met_avg
```

## 3. Use geom_violin to examine the wind speed and dew point temperature by region

#### Wind speed by region
Wind speed is lower in the eastern regions. The northeast region seems to have the largest range, as well as an outlier.
```{r violin-temp, cache=TRUE}
met_avg[!is.na(region)] %>% 
  ggplot(met_avg, mapping=aes(x=1, y=wind.sp)) + 
  geom_violin() + 
  facet_grid(~ region)
```

#### Dew point by region

The lowest dew points as well as the largest ranges in dew point are found in the western regions.
```{r violin-dew point, cache=TRUE}
met_avg[!is.na(region)] %>% 
  ggplot(met_avg, mapping=aes(x=1, y=dew.point)) + 
  geom_violin() + 
  facet_grid(~ region)
```

## 4. Use geom_jitter with stat_smooth to examine the association between dew point temperature and wind speed by region

The eastern regions seem to have a slightly positive association between dew point temperature and wind speed, as seen by the linear regression lines having a slightly positive slope. The linear regression lines for the western regions have slopes close to zero.

```{r geom_jitter, cache=TRUE}
met_avg[!is.na(dew.point) & !is.na(wind.sp)] %>% 
  ggplot(mapping=aes(x=dew.point, y = wind.sp, color=region, linetype=region)) + 
  geom_jitter() +
  stat_smooth(method=lm, se=FALSE, color="navy") + 
  scale_linetype_manual(values=c("solid","dotted", "twodash", "dashed"))
```

## 5. Use geom_bar to create barplots of the weather stations by elevation category coloured by region

The northeast has a similar number of weather stations in high and low elevations, as opposed to southeast region which has more weather stations at low elevations as well as the northwest and southwest regions which both have fewer weather stations at low elevations.
```{r geom_bar, cache=TRUE}
met_avg[!is.na(USAFID)] %>% 
  ggplot(mapping=aes(x=elev_cat, fill=region)) + 
  geom_bar(position="dodge") +
  labs(title = "Number of Weather Stations by Elevation Category and Region", x  = "Elevation Category", y = "Number of Weather Stations") +
  scale_fill_brewer(palette = "RdPu")
```

## 6. Use stat_summary to examine mean dew point and wind speed by region with standard deviation error bars

#### Mean Wind speed by region

The northeast and southeast regions have similar average wind speeds to each other, and the northwest and southwest regions have similar average wind speeds to each other. The eastern regions have lower average wind speeds compared to the western regions.
```{r stat_summary wind.sp, cache=TRUE}
ggplot(met_avg[!is.na(wind.sp)], mapping=aes(x=region, y=wind.sp)) +
  stat_summary(fun.data = "mean_sdl") + 
  stat_summary(fun.data = "mean_sdl", geom="errorbar")
```

#### Mean dew point by region

The northwest and southwest regions have similar average dew points. The southeast has the smallest standard error and highest average dew point. Average dew point temperature is higher in eastern regions than in western regions.
```{r, cache=TRUE}
ggplot(met_avg[!is.na(dew.point)], mapping=aes(x=region, y=dew.point)) +
  stat_summary(fun.data = "mean_sdl") + 
  stat_summary(fun.data = "mean_sdl", geom="errorbar")
```

## 7. Make a map showing the spatial trend in relative h in the US

Relative humidity temperature tends to increase as we move from western to eastern US.
```{r leaflet, cache=TRUE}
met_avg2 <- met_avg[!is.na(rh)]
temp.pal <- colorNumeric(c('darkgreen', 'goldenrod', 'brown'), domain = met_avg2$rh)

leaflet(met_avg2) %>%
  addProviderTiles('CartoDB.Positron') %>%
  addCircles(lat = ~lat, lng=~lon,
     label = ~paste0(round(rh,2), ' rh'), color = ~ temp.pal(rh),
     opacity = 1, fillOpacity = 1, radius = 500) %>%
  addMarkers(lat = ~met_avg2[rank(-rh) <= 10, lat], lng = ~met_avg2[rank(-rh) <= 10, lon]) %>%
  addLegend('bottomleft', pal=temp.pal, values=met_avg2$rh,
          title='Temperature, C', opacity=1)
  
```

## 8. Use a ggplot extension: ggpubr

ggpubr can be used to create publication-ready plots, including ones that have p-values included. This graph is a boxplot of relative humidity per region, with p values across groups added.
```{r ggpubr, cache=TRUE}
p <- ggboxplot(met_avg2, x = "region", y = "rh", color="region")
my_comparisons <- list( c("Southwest", "Northeast"),c("Northeast", "Southeast"), c("Southeast", "Northwest"), c("Southwest", "Southeast"), c("Northeast", "Northwest"))
p + stat_compare_means(comparisons = my_comparisons) # Add pairwise comparisons p-value
```