---
title: "Lab 3"
author: "Audrey Omidsalar"
date: "9/10/2021"
output:
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

## 1. Read in the Data

```{r}
download.file("https://raw.githubusercontent.com/USCbiostats/data-science-data/master/02_met/met_all.gz", "met_all.gz", method="libcurl", timeout = 60)
met <- data.table::fread("met_all.gz")

```

## 2. Check the dimensions, headers, footers. How many columns, rows are there?

There are 2377343 rows and 30 columns


```{r}
dim(met)
```
```{r}
head(met)
```
```{r}
tail(met)
```
## 3. Take a look at the variables
```{r}
str(met)
```
## 4. Take a closer look at the key variables
```{r}
table(met$year)
```
```{r}
table(met$day)
```
```{r}
table(met$hour)
```
```{r}
summary(met$temp)
```
```{r}
summary(met$elev)
```
```{r}
summary(met$wind.sp)
```
After checking the data we should make the appropriate modifications. Replace elevations with 9999 as NA.
```{r}
met[met$elev==9999.0] <- NA
summary(met$elev)
```
The weather station with highest elevation is 4113 meter (after replacing elevations of 9999 with NA)
```{r}
met <- met[temp>-40]
met2 <- met[order(temp)]
head(met2)
tail(met2)
```
## 5. Check the data against an external data source
```{r}
met <- met[temp>-15]
met2 <- met[order(temp)]
head(met2)
```
## 6. Calculate summary statistics
```{r}

```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
