---
title: "Lab 3"
author: "Audrey Omidsalar"
date: "9/10/2021"
output:
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library('leaflet')
library('lubridate')
library('data.table')
```

## R Markdown

## 1. Read in the Data

```{r}
if (!file.exists("met_all.gz")){
download.file("https://raw.githubusercontent.com/USCbiostats/data-science-data/master/02_met/met_all.gz", "met_all.gz", method="libcurl", timeout = 60)}
met <- data.table::fread("met_all.gz")
```

## 2. Check the dimensions, headers, footers. How many columns, rows are there?

There are 2377343 rows and 30 columns

```{r}
dim(met)
```
```{r}
head(met)
```
```{r}
tail(met)
```
## 3. Take a look at the variables
```{r}
str(met)
```
## 4. Take a closer look at the key variables
```{r}
table(met$year)
```
```{r}
table(met$day)
```
```{r}
table(met$hour)
```
```{r}
summary(met$temp)
```
```{r}
summary(met$elev)
```
```{r}
summary(met$wind.sp)
```
After checking the data we should make the appropriate modifications. Replace elevations with 9999 as NA.
```{r}
met[met$elev==9999.0] <- NA
summary(met$elev)
```
The weather station with highest elevation is 4113 meters (after replacing elevations of 9999 with NA)
```{r}
met <- met[temp>-40]
met2 <- met[order(temp)]
head(met2)
tail(met2)
```
## 5. Check the data against an external data source
Removed temperatures less than -15C, and made a new dataset (met2) which is ordered according to temperature. The new minimum temperature is -3C
```{r}
met <- met[temp>-15]
met2 <- met[order(temp)]
head(met2)
```
## 6. Calculate summary statistics
Select the weather station with maximum elevation
```{r}
elev <- met[elev==max(elev)]
summary(elev)
```
Correlation between temperature and wind speed
```{r}
cor(elev$temp, elev$wind.sp, use="complete")
```
Correlation between temperature and hour
```{r}
cor(elev$temp, elev$hour, use="complete")
```
Correlation between wind speed and day
```{r}
cor(elev$wind.sp, elev$day, use="complete")
```
Correlation between wind speed and hour
```{r}
cor(elev$wind.sp, elev$hour, use="complete")
```
Correlation between temperature and day
```{r}
cor(elev$temp, elev$day, use="complete")
```
## 7. Exploratory graphs
Looking at distribution of elevations
```{r}
hist(met$elev, breaks=100)
```
Distribution of temperature
```{r}
hist(met$temp)
```
Distribution of wind speed
```{r}
hist(met$wind.sp)
```
Mapping where the weather station with highest elevation is located.
```{r}
leaflet(elev) %>%
  addProviderTiles('OpenStreetMap') %>% 
  addCircles(lat=~lat,lng=~lon, opacity=1, fillOpacity=1, radius=100)
```
```{r}
library(lubridate)
elev$date <- with(elev, ymd_h(paste(year, month, day, hour, sep= ' ')))
summary(elev$date)
```
```{r}
elev <- elev[order(date)]
head(elev)
```
Time series plot - date versus temperature. The temperature fluctuates throughout the day, but the maximum and minimum temperatures are staying roughly the same throughout the month.
```{r}
plot(elev$date, elev$temp, type='l')
```
Time series plot - date versus wind speed. Wind speed fluctuates throughout the day, and peaks twice throughout the month (once around August 17th and again around August 25th).
```{r}
plot(elev$date, elev$wind.sp, type='l')
```