---
title: "Week 5 Lab"
author: "Audrey Omidsalar"
date: "9/24/2021"
output:
  html_document:
    toc: yes
    toc_float: yes
    keep_md: yes
  github_document:
  always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library('data.table')
library('dtplyr')
library('dplyr')
library('leaflet')
```

## Load Data

```{r load, cache=TRUE}
if (!file.exists("met_all.gz"))
  download.file(
    url = "https://raw.githubusercontent.com/USCbiostats/data-science-data/master/02_met/met_all.gz",
    destfile = "met_all.gz",
    method   = "libcurl",
    timeout  = 60
  )
met <- data.table::fread("met_all.gz")

#stations <- fread("ftp://ftp.ncdc.noaa.gov/pub/data/noaa/isd-history.csv")
stations <- fread("ftp://ftp.ncdc.noaa.gov/pub/data/noaa/isd-history.csv")
```

## Processing / Filtering Data

```{r filtering}
stations[, USAF := as.integer(USAF)]
stations[, USAF   := fifelse(USAF == 999999, NA_integer_, USAF)]
stations[, CTRY   := fifelse(CTRY == "", NA_character_, CTRY)]
stations[, STATE  := fifelse(STATE == "", NA_character_, STATE)]

# Selecting the three relevant columns, and keeping unique records
stations <- unique(stations[, list(USAF, CTRY, STATE)])
stations <- stations[!is.na(USAF)]
##easy way to remove duplicates
stations[, n := 1:.N, by = .(USAF)]
stations <- stations[n == 1,][, n := NULL]
```
### Merge the data
```{r merging}
merged <- merge(
  # Data
  x     = met,      
  y     = stations, 
  # List of variables to match
  by.x  = "USAFID",
  by.y  = "USAF", 
  # Which obs to keep?
  all.x = TRUE,      
  all.y = FALSE
)
```
## Question 1: Representative Station for the US
### What is the median station in terms of temperature, wind speed, and atmospheric pressure? Look for the three weather stations that best represent continental US using the quantile() function. Do these three coincide?

```{r question1}
##There are multiple measurements per station. We have to summarize the data first to characterize each station
merged_station_avg <- merged[,.(
  temp = mean(temp,na.rm=TRUE),
  wind.sp = mean(wind.sp, na.rm=TRUE),
  atm.press = mean(atm.press, na.rm=TRUE), 
  lon = mean(lon, na.rm = TRUE),
  lat = mean(lat, na.rm = TRUE)
), by= c("USAFID", "STATE")]
```
### Now find the quantiles per variable
```{r median}
merged_station_avg[, temp_med := quantile(temp, probs=0.5, na.rm=TRUE)]
merged_station_avg[, wind.sp_med := quantile(wind.sp, probs=0.5, na.rm=TRUE)]
merged_station_avg[, atm.press_med := quantile(atm.press, probs=0.5, na.rm=TRUE)]


medians <- merged_station_avg[,.(
  temp_50      = quantile(temp, probs=0.5, na.rm=TRUE),
  wind.sp_50   = quantile(wind.sp, probs=0.5, na.rm=TRUE),
  atm.press_50 = quantile(atm.press, probs=0.5, na.rm=TRUE)
)]
medians
```
### Find the station that is closest to these median values using which.min()
The median temperature station is at USAFID 720458.
The median wind speed station is at USAFID 720929.
The median atmospheric pressure station is at USAFID 722238.
The median stations for these three variables do not coincide.
```{r}
merged_station_avg[which.min(abs(temp - temp_med))]
merged_station_avg[which.min(abs(wind.sp - wind.sp_med))]
merged_station_avg[which.min(abs(atm.press - atm.press_med))]
#another method
#merged_station_avg[, temp_dist := abs(temp - medians$temp_50)]
#median_temp_station <- merged_station_avg[order(temp_dist)][1]
#median_temp_station
```
## Question 2 Representative Station per State
### Just like the previous question, you are asked to identify what is the most representative, the median, station per state. This time, instead of looking at one variable at a time, look at the euclidean distance. If multiple stations show in the median, select the one located at the lowest latitude.
```{r}
## add state column by merging
#merged_station_avg <- merge(x = merged_station_avg, y = stations, by.x = "USAFID", by.y = "USAF", all.x = TRUE, all.y=FALSE)
##Compute the median per state
merged_station_avg[,temp_med_state := quantile(temp, probs=.5, na.rm=TRUE), by = "STATE"]
merged_station_avg[,wind.sp_med_state := quantile(wind.sp, probs=.5, na.rm=TRUE), by = "STATE"]
merged_station_avg[,atm.press_med_state := quantile(atm.press, probs=.5, na.rm=TRUE), by = "STATE"]
```

```{r}
##Compute the Euclidean distance
summary(merged_station_avg$temp)
summary(merged_station_avg$wind.sp)
summary(merged_station_avg$atm.press)
##These datasets have NA's. Remove them to be able to calculate Euclidean distance.
merged_station_avg_noNA <- merged_station_avg[!is.na(temp) & !is.na(wind.sp) & !is.na(atm.press)]
##Calculate Euclidean Distance
merged_station_avg_noNA[, eudist := sqrt((temp - temp_med_state)^2 + (wind.sp - wind.sp_med_state)^2 + (atm.press - atm.press_med_state)^2)]
```
This `center_3vars` dataset shows the station with the minimum Euclidean distance (in terms of temperature, wind speed and atmospheric pressure) per state.
```{r}
center_3vars <- merged_station_avg_noNA[ , .SD[which.min(eudist)], by = STATE]
center_3vars
```
## Question 3: In the Middle?
### For each state, identify what is the station that is closest to the mid-point of the state. Combining these with the stations you identified in the previous question, use leaflet() to visualize all ~100 points in the same figure, applying different colors for those identified in this question.
```{r}
##Finding median latitude and longitude per state
merged[, lat_median := quantile(lat, probs=.5, na.rm=TRUE), by = "STATE"]
merged[, lon_median := quantile(lon, probs=.5, na.rm=TRUE), by = "STATE"]
##Finding the Euclidean distances between latitude and longitude
merged[, eudist_latlon := sqrt((lat - lat_median)^2 + (lon - lon_median)^2)]
##Outputting rows with minimum Euclidean distance per state
center_state <- merged[ , .SD[which.min(eudist_latlon)], by = STATE]
center_state
```
Combine the two datasets
```{r}
center_state[, label := "Center of the state"]
center_3vars[, label := "Center of wind speed, temperature, and atmospheric pressure"]
centers <- rbind(center_state, center_3vars, fill=TRUE)
```
Create map. The median center of state stations are shown in pink, and the stations corresponding to the median of the three variables (wind speed, atmospheric pressure and temperature) are in orange.
```{r}
leaflet(centers) %>%
  addProviderTiles("OpenStreetMap") %>%
  addCircles(lng = ~lon, lat = ~lat, color=~ifelse(label=="Center of the state",'pink','orange'), opacity=1,fillOpacity=0.7, radius=100)
```
## Question 4 - Mean of Means

```{r}
##creating average temperature per state
merged[, state_temp_mean := mean(temp, na.rm=TRUE), by = "STATE"]
##annotating temperature categories
merged[, temp_cat := fifelse(state_temp_mean < 20, "low",
                         fifelse(state_temp_mean < 25, "mid", "high"))]
table(merged$temp_cat, useNA = "always") ##will show number of NA's
```
Summary table
```{r}
##creating table with assigned variables
tab <- merged[, .(
  N_entries      = .N,
  N_na           = sum(is.na(temp)),
  N_stations     = length(unique(USAFID)),
  N_states       = length(unique(STATE)),
  mean_temp      = mean(temp, na.rm=TRUE),
  mean_wind.sp   = mean(wind.sp, na.rm=TRUE),
  mean_atm.press = mean(atm.press, na.rm=TRUE)
)
    , by = "temp_cat"]
##make the table look pretty
knitr::kable(tab)
```

